\documentclass{article}
\usepackage[UTF8]{ctex}
\usepackage{amsmath} % 使用 amsmath 宏包以改进公式排版
% !TEX TS-program = xelatex
% !TEX encoding = UTF-8 Unicode
% !Mode:: "TeX:UTF-8"


\setmainfont{Times New Roman}
\usepackage{geometry}
\geometry{a4paper,scale=0.8}
\usepackage{algorithm} % 引入 algorithm 宏包
\usepackage{algpseudocode} % 引入 algpseudocode 用于编写伪代码
\usepackage{amsthm}

\begin{document}
%\{\heiti 黑体}，{\kaishu 楷体} 

\begin{center}
{\huge \textbf{Exercis 1}}

\textit{Exercises for Algorithms by Nengjun Zhu.}

\vspace{1em}
\textbf{Name:} \underline{\qquad \textbf{郭同} \qquad} \,
\textbf{Student ID:}\underline{\qquad \textbf{24721537}\qquad}\, 
\textbf{Email:}\underline{\qquad \textbf{teri@shu.edu.cn}\qquad}
\end{center}

\begin{itemize}

\item[1.] Suppose $a_0=1,  a_1=2, a_2=3, a_k=a_{k-1} + a_{k-2} + a_{k-3} $ for $k\geq 3$. Use strong principle of mathematical induction to prove that $a_n \leq 2^n$ for all integers $n\geq0$.





\noindent 

\paragraph{证:  } 当 $0 \leq n \leq 4$ 时，结论显然成立。

\paragraph{假设} 对于 $5 \leq k$，假设有 $a_k \leq 2^k$，则我们有：
\begin{align*}
    a_{k+1} &= a_k + a_{k-1} + a_{k-2} \\
            &\leq 2^k + 2^{k-1} + 2^{k-2} \\
            &= 2^{k-2} (4 + 2 + 1) \\
            &\leq 2^{k+1}
\end{align*}

\paragraph{结论} 因此结论成立。






\item[2.] {\large Consider the sorting algorithm shown in Alg.1, which is called BUBBLESORT.}
\begin{itemize}
\item[(a)] What is the minimum number of element comparisons? When is this minimum achieved?

\paragraph{答:}当元素为升序时比较次数最少,为$n - 1$ 次

\item[(b)] What is the maximum number of element comparisons? When is this maximum achieved?

\paragraph{答:}当元素为降序时比较次数最多,为$ \frac{n(n+1)}{2} $ 次

\item[(c)] Express the running time of Alg.1 in terms of the Ο and Ω notations.

\paragraph{答:} 令$g(n) = n - 1$,$h(n) =  \frac{n(n+1)}{2} $ ,可得$f(n) = O(h(n))$, $f(n) = \Omega(g(n))$
,时间复杂度为$O(n^2)$,
$\Omega(n)$。
\item[(d)] Can the running time of the algorithm be expressed in terms of the Θ notation? Explain.

\paragraph{答:} 平均情况下，数组元素是随机排列的，需要进行大约一半的交换操作，时间复杂度依然是$Θ(n^2)$。无论是平均情况还是最坏情况，冒泡排序的时间复杂度都是$Θ(n^2)$。因此，Θ表示法是合适的。
\end{itemize}



\item[3.] Fill in the blanks with either true (T) or false (F):

  \begin{center}
    \begin{tabular}{|c|c|c|c|c|} 
    \hline
    $f(n)$ & $g(n)$ & $f=O(g) $ &$f=\Omega(g)$  &$f=\Theta(g)$ \\
    \hline\hline
    $2n^3+3n$ & $100n^2 + 2n +100$ &F &T&F \\\hline
    $50n + \mbox{log}n $ & $10n+\mbox{log}\mbox{log}n$ &T &T& T\\\hline
    $50n\mbox{log}n$ & $10n\mbox{log}\mbox{log}n$ & F&T& F\\\hline
    $\mbox{log}n$ & $\mbox{log}^2n$ &T &F&F \\\hline
    $n!$ & $5^n$ &F  &T&F\\
    \hline
    \end{tabular}
  \end{center}


\item[4.] Design a divide-and-conquer algorithm to determine whether two given binary trees $T_1$ and $T_2$ are identical.

\begin{center}
\begin{algorithm}
  \caption{isSameTree: Check if two binary trees are identical}
  \begin{algorithmic}[1] % [1] 启用行号
      \State \textbf{Input:} Two binary trees, \textit{root1} and \textit{root2}
      \State \textbf{Output:} Boolean value indicating whether the trees are identical
  
      \Function{isSameTree}{root1, root2} % 定义函数
          \If{root1 = NULL and root2 = NULL} % 开始条件判断
              \State \Return true
          \EndIf % 正确结束 If
          \If{root1 = NULL or root2 = NULL}
              \State \Return false
          \EndIf
          \If{root1.val $\neq$ root2.val}
              \State \Return false
          \EndIf
          \State \Return \Call{isSameTree}{root1.left, root2.left} and \Call{isSameTree}{root1.right, root2.right}
      \EndFunction % 确保函数有结束
  \end{algorithmic}
  \end{algorithm}

\end{center}
  \paragraph{ }

\item[5.] You are given two sorted lists of size $m$ and $n$ in ascending order. Given an $O(\mbox{log}m + \mbox{log}n)$ time algorithm for computing the $k$-th smallest element in the union of the two lists.




\section*{Algorithm Idea}

1. \textbf{Base Case Handling}:
   \begin{itemize}
       \item If one of the lists is empty, the $k$-th smallest element is simply the $k$-th element of the other list.
       \item If $k = 1$, return the smaller of the first elements of the two lists.
   \end{itemize}

2. \textbf{Recursive Strategy}:
   \begin{itemize}
       \item Assume the length of list $A$ is always less than or equal to the length of list $B$ (swap if necessary). This ensures the first list is never longer than the second, optimizing recursive calls.
       \item Choose $i = \min(\frac{k}{2}, m)$ and $j = \min(\frac{k}{2}, n)$ as the midpoints to split the lists $A$ and $B$.
       \item Compare $A[i-1]$ and $B[j-1]$:
       \begin{itemize}
           \item If $A[i-1] < B[j-1]$, it means that the $k$-th smallest element cannot be in the first $i$ elements of $A$. Recursively search for the $(k - i)$-th smallest element in the remaining part of $A$ and the entire $B$.
           \item Otherwise, the $k$-th smallest element cannot be in the first $j$ elements of $B$. Recursively search for the $(k - j)$-th smallest element in $A$ and the remaining part of $B$.
       \end{itemize}
   \end{itemize}

3. \textbf{Time Complexity}:
   \begin{itemize}
       \item The algorithm halves the search space in each recursive step, achieving a time complexity of $O(\log m + \log n)$.
   \end{itemize}

\section*{Detailed Steps}
\begin{enumerate}
    \item \textbf{Input:} Two sorted lists $A$ of size $m$ and $B$ of size $n$, and an integer $k$.
    \item \textbf{Output:} The $k$-th smallest element in the union of $A$ and $B$.
    \item \textbf{Procedure:}
    \begin{enumerate}
        \item If $m > n$, swap $A$ and $B$ to ensure $A$ is the smaller list.
        \item If $m = 0$, return $B[k-1]$.
        \item If $k = 1$, return $\min(A[0], B[0])$.
        \item Set $i = \min(\frac{k}{2}, m)$ and $j = \min(\frac{k}{2}, n)$.
        \item Compare $A[i-1]$ and $B[j-1]$:
        \begin{itemize}
            \item If $A[i-1] < B[j-1]$, recursively find the $(k - i)$-th smallest element in $A[i:]$ and $B$.
            \item If $A[i-1] \geq B[j-1]$, recursively find the $(k - j)$-th smallest element in $A$ and $B[j:]$.
        \end{itemize}
    \end{enumerate}
\end{enumerate}


\end{itemize}
\end{document}